Demonstrating the Reliability of Randomized Measurement and Verification for Switchable Control Retrofits Using a Large Open-Source Dataset

### 2025-07-06 code/manuscript.Rmd
temp commit for figure update

---
title: "Demonstrating the Reliability of Randomized Measurement and Verification for Switchable Control Retrofits Using a Large Open-Source Dataset"
author:
  - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA], Paul Raftery ^1^, Stefano Schiavon ^1^, Carlos Duarte ^1^, Gail Brager ^1^
abstract: Conventional measurement and verification (M&V) methods rely on comparing pre- and post-retrofit performance to estimate energy savings. They are often time-consuming and unreliable, especially when non-routine events (step changes or more gradual changes in how a building operates) occur during the M&V process. Those events are unrelated to the intervention but significantly affect building energy consumption and thus when the analyst applies the conventional M&V method, the results can be confounded. In this study, we demonstrated that switchable interventions, such as most control retrofits, can benefit from a new M&V method that randomly samples whether to implement the baseline or the intervention strategy at a fixed interval (e.g., daily). We tested this novel randomized M&V method on a large public dataset (639 buildings) covering various climate zones and commercial building types, using a virtual chilled water supply temperature reset based on outdoor weather as the intervention. The results show that compared to the conventional method, the randomized method provides more accurate savings estimations with a median of 74% accuracy improvement and is even faster (typically 36 weeks instead of 2 years). Additionally, we found that when non-routine events are present, the randomized method estimates savings that are much closer to the ground-truth savings than the conventional method, demonstrating much improved reliability. We also assessed the impact of normalizing for different weather, starting the M&V at different dates of the year, continuing randomization with a different sampling ratio after satisfying all criteria, and dropping samples affected by carryover effects when switching between strategies. For each scenario, we identified the optimal sampling interval using the large dataset.

@@

# define base ggplot theme
theme_update(plot.title = element_text(size = 14, colour = "grey20", face = "bold", hjust = 0.5),
             plot.subtitle = element_text(size = 10, colour = "grey20", face = "italic", hjust = 0.5, margin = margin(b = 10)),
             plot.caption = element_text(size = 10, colour = "grey20", face = "italic", hjust = 0.5),
             plot.background = element_rect(fill = "white", colour = NA),
             panel.grid.minor = element_blank(),
             panel.grid.major = element_blank(),
             axis.text = element_text(size = 10),
             strip.text = element_text(size = 10, color = "grey20", face = "bold"),
             strip.background = element_blank())

@@

```{r graphicalabstract, out.width = "80%"}

knitr::include_graphics("../figs/manuscript/graphical_abs.jpg", dpi = 900)
```

@@

-   Verified the improvements of the randomized switchback method using
    a large open-source dataset
-   Addressed the influence of non-routine events and also more subtle
    long-term 'operational drift' in buildings
-   Investigated the trade-off of selecting longer sampling intervals to
    account for carryover effect and weather dependency of the method
-   Discussed the impact of selecting different sampling ratio on the
    M&V accuracy

@@

To address the limitations of conventional M&V methods and the
challenges posed by building baseline changes, we proposed a M&V method
that adopts the randomized crossover design [@raftery2024], a gold
standard from medical, agricultural and online controlled experimental
(i.e. A/B testing) studies [@blackston2019; @duan2013; @gupta2019;
@mie2022; @munro; @raseduzzaman2017]. An early application of this
randomized design to the field of M&V in buildings was first used in
[@raftery2018], which compared the energy performance of two different
supply air temperature reset strategies in a building. Subsequent
improvements to the M&V method include adding a sequential evaluation
framework and defining stopping criteria to end the M&V period early
after reaching a target level of uncertainty, using blocked
randomization by weekday. The full framework is detailed in a previously
published study [@raftery2024] along with a case study application to a
real building, and other example use cases. In summary, this method
provides M&V analysts with a randomized schedule that alternates between
baseline and intervention implementation while ensuring balanced
sampling across days of the week and seasons. For example, given a
10-week M&V period for one intervention, the balanced randomized
schedule would equally sample five Mondays with the baseline strategy in
operation and five Mondays with the intervention strategy in operation.
The major limitation of this method is that it is only applicable to a
subset of retrofit projects where interventions can be easily switched
on and off, which is a small subset of all the interventions performed
in buildings. However, for all applicable use cases, it allows analysts
to detect energy savings sequentially shortly after the test begins. In
addition, once the desired estimate uncertainty target is achieved,
analysts can terminate the M&V study and switch to 100% intervention.
The key advantage of randomization is that if control strategies are
sampled with equal probability, the influence of other confounding
factors such as occupancy change (a common non-routine event) and other
more subtle long term changes in building energy use are likely to be
evenly distributed among measurements, leading to a more accurate and
unbiased assessment of the intervention effect.

@@

```{r stablesite, fig.cap = "Site summary of the filtered buildings from BDG2 dataset (counts < 10 are omitted for visualization; left: aggregated counts of buildings for each type; right: breakdown building counts for each building type at each location)", fig.width=8, fig.height=6}

set3 <- colorRampPalette(brewer.pal('Set3',n=12))
type_colors <- setNames(set3(13), all_types_variable$type)

stable_set <- df_energy_stable %>%
  group_by(type) %>%
  distinct(name) %>%
  summarise(stable = n()) %>%
  ungroup()

variable_set <- df_energy_variable %>%
  group_by(type) %>%
  distinct(name) %>%
  summarise(variable = n()) %>%
  ungroup()

p1 <- variable_set %>% 
  left_join(stable_set, by = "type") %>% 
  mutate(stable = replace_na(stable, 0)) %>% 
  group_by(type) %>% 
  summarise(n = stable + variable) %>% 
  ungroup() %>% 
  mutate(type = fct_reorder(type, n, .desc = F)) %>%
  ggplot(aes(x = 1, y = n, fill = as.factor(type))) +
  geom_col() +
  geom_text(aes(label = ifelse(n > 10, n, "")), color = "black", position = position_stack(vjust = 0.5)) +
  scale_y_continuous(expand = c(0, 0),
                     breaks = breaks_pretty(n = 4)) +
  scale_x_discrete(expand = c(0, 0.1)) +
  scale_fill_manual(values = type_colors) +
  labs(x = NULL,
       y = NULL,
       subtitle = "Combining all locations",
       fill = NULL) +
  theme(axis.text.x = element_blank(),
        panel.grid.major.y = element_line(color = "grey80"),
        legend.direction = "horizontal",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

stable_set <- df_energy_stable %>% 
  left_join(df_loc, by = "site") %>% 
  group_by(type, location) %>%
  distinct(name) %>%
  summarise(stable = n()) %>%
  ungroup()

variable_set <- df_energy_variable %>% 
   left_join(df_loc, by = "site") %>% 
  group_by(type, location) %>%
  distinct(name) %>%
  summarise(variable = n()) %>%
  ungroup()

p2 <- variable_set %>% 
  left_join(stable_set, by = c("type", "location")) %>% 
  mutate(stable = replace_na(stable, 0)) %>% 
  group_by(location, type) %>% 
  summarise(n = stable + variable) %>% 
  ungroup() %>% 
  group_by(location) %>%
  mutate(proportion = n / sum(n),
         total = sum(n),
         ymax = cumsum(proportion),
         ymin = c(0, head(ymax, n = -1))) %>%
  mutate(label_pos = (ymax + ymin) / 2) %>%
  ungroup() %>%
  group_by(type) %>%
  mutate(order = sum(n)) %>%
  ungroup() %>%
  mutate(type = fct_reorder(type, order, .desc = F)) %>%
  ggplot(aes(ymax = ymax, ymin = ymin, xmax = 4, xmin = 3.2, fill = type)) +
  geom_rect() +
  coord_polar(theta = "y") +
  xlim(c(2, 4)) +
  facet_wrap(~ location, nrow = 3) +
  labs(x = NULL,
       y = NULL,
       subtitle = "For each location",
       fill = NULL) +
  scale_fill_manual(values = type_colors) +
  geom_text(aes(x = 3.7, y = label_pos, label = ifelse(n > 10, n, ""))) +
  geom_text(aes(x = 2, y = 0, label = paste0("Total\n", total)), color = "black", check_overlap = T) +
  theme(legend.direction = "horizontal",
        axis.text = element_blank(),
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

ggarrange(p1, p2,
          ncol = 2, nrow = 1,
          widths = c(0.2, 1),
          common.legend = TRUE,
          legend = "bottom") +
  plot_annotation(title = "Summary of filtered buildings and types")
```

@@

data.frame(
  OAT = c(10, 15, 25, 30),  
  Baseline = c(6, 6, 6, 6),      
  Intervention = c(12, 12, 7, 7)
  ) %>% 
  pivot_longer(cols = c("Baseline", "Intervention"), names_to = "strategy", values_to = "value") %>% 
  ggplot(aes(x = OAT, y = value, color = strategy)) +
  geom_line(linewidth = 1.2) +
  scale_color_manual(values = ls_colors) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = c(4, 6, 8, 10, 12),
                     limits = c(5, 12.5),
                     labels = number_format(suffix = " °C")) +
  scale_x_continuous(expand = c(0, 0), 
                     breaks = c(5, 10, 15, 20, 25, 30), 
                     labels = number_format(suffix = " °C")) +
  geom_vline(xintercept = 10, lty = "dashed", color = "red") +
  annotate(geom = "text", x = 7.5, y = 8, label = "No chiller operation\nbelow 10 °C") +
  coord_cartesian(xlim = c(5, 30)) +
  labs(x = "Outdoor temperature", 
       y = "Chilled water supply temperature", 
       color = NULL, 
       title = "Chilled water supply temperature reset intervention") +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

@@

| Parameter              | Description                                                      | Value |
|-------------------|-----------------------------------|-------------------|
| $\beta_{temp}$         | \% savings from setting $T_{interv}$ 1 °C higher than $T_{base}$ | 0.08  |
| $\beta_{occ}$          | \% savings adjustment during occupied hours                      | 1.2   |
| $\beta_{\text{unocc}}$ | \% savings adjustment during unoccupied hours                    | 0.8   |

@@

```{r abstract, fig.cap="Overall comparison results between the conventional M&V method and the proposed randomized M&V method (both upon reaching the stopping criteria for each case, and over a fixed two-year period).", fig.height=11, fig.width=12}

@@

p_top <- df_MW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
  geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
            aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
            size = 4) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       color = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with measured weather")) +
  coord_cartesian(ylim = c(0, 23)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        axis.text.x = element_blank(),
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

@@

p_middle <- df_TW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
  geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
            aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
            size = 4) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
  coord_cartesian(ylim = c(0, 23)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

@@

p_timeline <- count %>% 
  ggplot() +
  geom_bar(data = . %>% 
             filter(n_weeks == 24 | n_weeks == 36),
           aes(x = n_weeks, y = eob, fill = "Buildings finishing randomized M&V"), 
           stat = "identity", 
           position = "stack",
           alpha = 0.8, 
           width = 3) +
  geom_line(aes(x = n_weeks, y = temp, color = "Buildings satisfying 80% TMY range"),
            alpha = 0.2) +
  geom_point(aes(x = n_weeks, y = temp, color = "Buildings satisfying 80% TMY range"), 
             size = 1.5) +
  geom_line(aes(x = n_weeks, y = sprt, color = "Buildings satisfying SPRT"),
            alpha = 0.2) +
  geom_point(aes(x = n_weeks, y = sprt, color = "Buildings satisfying SPRT"), 
             size = 1.5, 
             shape = 17) +
  geom_segment(aes(x = 36.5, y = max(eob), xend = 95.5, yend = max(eob)),
               arrow = arrow(length = unit(0.25, "in")),   
               linewidth = 1.1,  
               color = "#fb8072") + 
  annotate(geom = "text", 
           x = 66, 
           y = max(count$eob) - 30, 
           size = 5,
           label = "Excess time by conventional M&V") +
  geom_vline(xintercept = c(12, 24, 36, 48, 96), lty = "dashed", color = "grey80") +
  annotate(geom = "text", 
           x = seq(6, 48, by = 12), 
           y = 200, 
           label = paste0("12-week\nblock"), 
           alpha = 0.5) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = ls_colors) +
  scale_x_continuous(expand = c(0, 0), 
                     limits = c(0, 100),
                     breaks = c(12, 24, 36, 48, 96), 
                     labels = c("12 weeks", "24 weeks", "36 weeks", "1 year", "2 years")) +
  coord_cartesian(ylim = c(0, 650)) +
  labs(x = NULL, 
       y = "Number of buildings", 
       fill = NULL, 
       color = NULL, 
       subtitle = "timeline comparison") +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

ggarrange(p_accuracy, p_timeline, 
          ncol = 1, nrow = 2,
          heights = c(2.2, 1),
          labels = c("", "c)"),
          common.legend = F, 
          legend = "bottom") +
  plot_annotation(title = "Overall accuracy and time required for savings estimation by different M&V methods")

@@

```{r meanerr, fig.cap="Distribution of estimated deviation from the ground-truth savings calculated by the two M&V methods (red dashed line highlighting 0% error)", fig.height=8, fig.width=12}
# plot for the stable subset
rand_eob_S <- df_seq_fs_stable %>% 
  filter(seq == "eob") %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = savings - fs, 
         method = "rand_eob") %>% 
  select(name, method, diff)

@@

p_top <- df_MW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
  geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
            aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
            size = 4) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       color = NULL, 
       x = NULL, 
       y = "Error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with measured weather")) +
  coord_cartesian(ylim = c(-20, 20)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        axis.text.x = element_blank(),
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

@@

p_middle <- df_TW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
  geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
            aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
            size = 4) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       x = NULL, 
       y = "Estimated error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
  coord_cartesian(ylim = c(-20, 20)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

@@

```{r nosaving, fig.cap="Comparison between the two M&V methods in detecting no intervention effect when buildings are subject to static non-rountine events and gradual operational drift.", fig.height=5, fig.width=12}

>>>>

---
title: "Demonstrating the Reliability of Randomized Measurement and Verification for Switchable Control Retrofits Using a Large Open-Source Dataset"
author:
  - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA], Paul Raftery ^1^, Stefano Schiavon ^1^ ^[Department of Architecture, College of Environmental Design, University of California, Berkeley, USA] ^[Department of Civil and Environmental Engineering, College of Engineering, University of California, Berkeley, USA], Carlos Duarte ^1^, Gail Brager ^1^
abstract: Conventional measurement and verification (M&V) methods rely on comparing pre- and post-retrofit performance to estimate energy savings. They are often time-consuming and unreliable, especially when non-routine events, as step changes or more gradual changes in how a building operates, occur during the M&V process. Those events are unrelated to the intervention but significantly affect building energy consumption and thus when the analyst applies the conventional M&V method, the results can be confounded. In this study, we demonstrated that switchable interventions, such as most HVAC control retrofits, can benefit from a new M&V method that randomly samples whether to implement the baseline or the intervention strategy at a fixed interval (e.g., daily). We tested this novel randomized M&V method on a large public dataset (639 buildings) covering various climate zones and commercial building types, using a virtual chilled water supply temperature reset based on outdoor weather as the intervention. The results show that compared to the conventional method, the randomized method provides more accurate savings estimations with a median of 74% accuracy improvement and is even faster (typically 36 weeks instead of 104 weeks, ~65% reduction). Additionally, we found that when non-routine events are present, the randomized method estimates savings that are much closer to the ground-truth savings than the conventional method, demonstrating much improved reliability. We also assessed the impact of normalizing for different weather, starting the M&V at different dates of the year, continuing randomization with a different sampling ratio after satisfying all criteria, and dropping samples affected by carryover effects when switching between strategies. For each scenario, we identified the optimal sampling interval using the large dataset.

@@

# define base ggplot theme
theme_update(plot.title = element_text(size = 16, colour = "grey20", face = "bold", hjust = 0.5),
             plot.subtitle = element_text(size = 14, colour = "grey20", face = "italic", hjust = 0.5, margin = margin(b = 10)),
             plot.caption = element_text(size = 10, colour = "grey20", face = "italic", hjust = 0.5),
             plot.background = element_rect(fill = "white", colour = NA),
             panel.grid.minor = element_blank(),
             panel.grid.major = element_blank(),
             axis.text = element_text(size = 12),
             legend.text = element_text(size = 12),
             strip.text = element_text(size = 12, color = "grey20", face = "bold"),
             strip.background = element_blank())

@@

```{r graphicalabstract, out.width = "100%"}

knitr::include_graphics(paste0(fig_path, "graphical_abs.jpg"), dpi = 900)
```

@@

-   Randomized switchback method leads to time reduction and increased
    accuracy
-   Addressed the influence of non-routine events and also more subtle
    long-term 'operational drift' in buildings
-   Investigated the trade-off of selecting longer sampling intervals to
    account for carryover effect and weather dependency of the method
-   Discussed the impact of selecting different sampling ratio on the
    M&V accuracy

@@

To address the limitations of conventional M&V methods and the
challenges posed by building baseline changes, we proposed a M&V method
that adopts the randomized crossover design [@raftery2024], a gold
standard from medical, agricultural and online controlled experimental
(i.e. A/B testing) studies [@blackston_comparison_2019; @duan2013;
@gupta2019; @mie2022; @munro; @raseduzzaman2017]. An early application
of this randomized design to the field of M&V in buildings was first
used in [@raftery2018], which compared the energy performance of two
different supply air temperature reset strategies in a building.
Subsequent improvements to the M&V method include adding a sequential
evaluation framework and defining stopping criteria to end the M&V
period early after reaching a target level of uncertainty, using blocked
randomization by weekday. The full framework is detailed in a previously
published study [@raftery2024] along with a case study application to a
real building, and other example use cases. In summary, this method
provides M&V analysts with a randomized schedule that alternates between
baseline and intervention implementation while ensuring balanced
sampling across days of the week and seasons. For example, given a
10-week M&V period for one intervention, the balanced randomized
schedule would equally sample five Mondays with the baseline strategy in
operation and five Mondays with the intervention strategy in operation.
The major limitation of this method is that it is only applicable to a
subset of retrofit projects where interventions can be easily switched
on and off, which is a small subset of all the interventions performed
in buildings. However, for all applicable use cases, it allows analysts
to detect energy savings sequentially shortly after the test begins. In
addition, once the desired estimate uncertainty target is achieved,
analysts can terminate the M&V study and switch to 100% intervention.
The key advantage of randomization is that if control strategies are
sampled with equal probability, the influence of other confounding
factors such as occupancy change (a common non-routine event) and other
more subtle long term changes in building energy use are likely to be
evenly distributed among measurements, leading to a more accurate and
unbiased assessment of the intervention effect.

@@

@@

data.frame(
  OAT = c(10, 15, 25, 30),  
  Baseline = c(6, 6, 6, 6),      
  Intervention = c(12, 12, 7, 7)
  ) %>% 
  pivot_longer(cols = c("Baseline", "Intervention"), names_to = "strategy", values_to = "value") %>% 
  ggplot(aes(x = OAT, y = value, color = strategy)) +
  geom_line(linewidth = 1.2) +
  scale_color_manual(values = ls_colors) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = c(4, 6, 8, 10, 12),
                     limits = c(5, 12.5),
                     labels = number_format(suffix = " °C")) +
  scale_x_continuous(expand = c(0, 0), 
                     breaks = c(5, 10, 15, 20, 25, 30), 
                     labels = number_format(suffix = " °C")) +
  geom_vline(xintercept = 10, lty = "dashed", color = "red") +
  annotate(geom = "text", x = 7.5, y = 8, label = "No chiller\noperation\nbelow 10 °C", size = 5) +
  coord_cartesian(xlim = c(5, 30)) +
  labs(x = "Outdoor temperature", 
       y = "Chilled water supply temperature", 
       color = NULL, 
       title = "Chilled water supply temperature reset intervention") +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

@@

| Parameter              | Description                                                      | Value |
|----------------|-----------------------------------------|----------------|
| $\beta_{temp}$         | \% savings from setting $T_{interv}$ 1 °C higher than $T_{base}$ | 0.08  |
| $\beta_{occ}$          | \% savings adjustment during occupied hours                      | 1.2   |
| $\beta_{\text{unocc}}$ | \% savings adjustment during unoccupied hours                    | 0.8   |

@@

```{r abstract, fig.cap="Overall comparison results between the conventional M&V method and the proposed randomized M&V method (both upon reaching the stopping criteria for each case, and over a fixed two-year period).", fig.height=10, fig.width=9}

@@

p_top <- df_MW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method), linewidth = 0.25) +
  geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
            aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
            size = 5) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       color = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with measured weather")) +
  coord_cartesian(ylim = c(0, 23)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        axis.text.x = element_blank(),
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

@@

p_middle <- df_TW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method), linewidth = 0.25) +
  geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
            aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
            size = 5) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
  coord_cartesian(ylim = c(0, 23)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

@@

p_timeline <- count %>% 
  ggplot() +
  geom_bar(data = . %>% 
             filter(n_weeks == 24 | n_weeks == 36),
           aes(x = n_weeks, y = eob, fill = "Buildings finishing randomized M&V"), 
           stat = "identity", 
           position = "stack",
           alpha = 0.8, 
           width = 3) +
  geom_line(aes(x = n_weeks, y = temp, color = "Buildings satisfying 80% TMY range"),
            alpha = 0.2) +
  geom_point(aes(x = n_weeks, y = temp, color = "Buildings satisfying 80% TMY range"), 
             size = 1.5) +
  geom_line(aes(x = n_weeks, y = sprt, color = "Buildings satisfying SPRT"),
            alpha = 0.2) +
  geom_point(aes(x = n_weeks, y = sprt, color = "Buildings satisfying SPRT"), 
             size = 1.5, 
             shape = 17) +
  geom_segment(aes(x = 36.5, y = max(eob), xend = 95.5, yend = max(eob)),
               arrow = arrow(length = unit(0.25, "in")),   
               linewidth = 1.1,  
               color = "#fb8072") + 
  annotate(geom = "text", 
           x = 66, 
           y = max(count$eob) - 35, 
           size = 5,
           label = "Excess time by conventional M&V") +
  geom_vline(xintercept = c(12, 24, 36, 48, 96), lty = "dashed", color = "grey80") +
  annotate(geom = "text", 
           x = seq(6, 48, by = 12), 
           y = 200, 
           size = 4, 
           label = paste0("12-week\nblock"), 
           alpha = 0.5) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = ls_colors) +
  scale_x_continuous(expand = c(0, 0), 
                     limits = c(0, 100),
                     breaks = c(12, 24, 36, 48, 96), 
                     labels = c("12 weeks", "24 weeks", "36 weeks", "1 year", "2 years")) +
  coord_cartesian(ylim = c(0, 650)) +
  labs(x = NULL, 
       y = "Number of buildings", 
       fill = NULL, 
       color = NULL, 
       subtitle = "timeline comparison") +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

ggarrange(p_accuracy, p_timeline, 
          ncol = 1, nrow = 2,
          heights = c(2.2, 1),
          labels = c("", "c)"),
          common.legend = F, 
          legend = "bottom") +
  plot_annotation(title = "Overall savings estimation comparison by different M&V methods")

@@

```{r meanerr, fig.cap="Distribution of estimated deviation from the ground-truth savings calculated by the two M&V methods (red dashed line highlighting 0% error)", fig.height=7, fig.width=9}
# plot for the stable subset
rand_eob_S <- df_seq_fs_stable %>% 
  filter(seq == "eob") %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = savings - fs, 
         method = "rand_eob") %>% 
  select(name, method, diff)

@@

p_top <- df_MW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method), linewidth = 0.25) +
  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
  geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
            aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
            size = 5) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 5), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       color = NULL, 
       x = NULL, 
       y = "Estimated error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with measured weather")) +
  coord_cartesian(ylim = c(-15, 15)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        axis.text.x = element_blank(),
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

@@

p_middle <- df_TW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method), linewidth = 0.25) +
  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
  geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
            aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
            size = 5) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 5), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       x = NULL, 
       y = "Estimated error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
  coord_cartesian(ylim = c(-15, 15)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

@@

```{r nosaving, fig.cap="Comparison between the two M&V methods in detecting no intervention effect when buildings are subject to static non-rountine events and gradual operational drift.", fig.height=4, fig.width=9}
